---
title: "R Notebook"
output: html_notebook
---

USAGE: variables generated with "kod_real.Rmd" script are needed to properly clip the area of interest

SETTING WORKING DIRECTORY
```{r}
setwd("D:/AGH/INZ/data")
```

LOADING PACKAGES
```{r}
library(ncdf4)
library(sf)
library(zoo)
library(chron)
library(RColorBrewer)
library(lattice)
library(weathermetrics)
```

##################################################################################################################################################

# FORECAST

Loading data
```{r}
ncin_fore_ppt<-nc_open("forecast/CHELSAcmip5ts_pr_ACCESS1-3_rcp45_2006-2029_V1.1.nc")
ncin_fore_tmax<-nc_open("forecast/CHELSAcmip5ts_tasmax_ACCESS1-3_rcp45_2006-2029_V1.1.nc")
ncin_fore_tmin<-nc_open("forecast/CHELSAcmip5ts_tasmin_ACCESS1-3_rcp45_2006-2029_V1.1.nc")

```

# MAX TEMPERATURE

Data details
```{r}
#print(ncin_fore_tmax)

#air_temperature[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

ncatt_get(ncin_fore_tmax, 0, "Title")
ncatt_get(ncin_fore_tmax, 0, "References")
ncatt_get(ncin_fore_tmax, 0, "history")
ncatt_get(ncin_fore_tmax, 0, "Conventions")

#Checking units
ncatt_get(ncin_fore_tmax,"longitude","units") #degrees_east
ncatt_get(ncin_fore_tmax,"latitude","units") #degrees_north
ncatt_get(ncin_fore_tmax,"time","units") # "days since 2006-1-15 00:00:00"
ncatt_get(ncin_fore_tmax,"air_temperature","units") #K

ncatt_get(ncin_fore_tmax,"air_temperature","long_name") #Daily Maximum Near-Surface Air Temperature
ncatt_get(ncin_fore_tmax,"air_temperature","_FillValue") #-99999
ncatt_get(ncin_fore_tmax,"air_temperature","missing_value") #-99999


```



Extracting variables
```{r}
lon_fore <- ncvar_get(ncin_fore_tmax, "longitude")
lat_fore <- ncvar_get(ncin_fore_tmax, "latitude")
time_fore <- ncvar_get(ncin_fore_tmax, "time")

dim(lon_fore) #8017
dim(lat_fore) #3875
dim(time_fore) # 288 -> 24 years, 12 months each

head(lon_fore) # - => +
head(lat_fore) # + => -
head(time_fore)

```


Time units correction
```{r}
time_units <- ncatt_get(ncin_fore_tmax,"time","units")
time_units <- strsplit(time_units$value, " ")
time_units <- strsplit(unlist(time_units)[3], "-")

time_year <- as.integer(unlist(time_units)[1])
time_month <- as.integer(unlist(time_units)[2])
time_day <- as.integer(unlist(time_units)[3])

time_fore <- chron(time_fore, origin = c(time_month, time_day, time_year))
Sys.setlocale("LC_TIME", "C")
time_fore <- as.yearmon(time_fore)

head(time_fore)
tail(time_fore)
#time range: Jan 2006 - Dec 2029

#limit to range Jan 2006 - Dec 2021
time_fore <- time_fore[1:192]
#which(time_fore == tail(time_fore, n=1))


#extracting values only from July
#july <- seq(7, 12*16, by=12)
#july
#time_fore_july <- time_fore[july]
#time_fore_july

#length(time_fore_july) == length(time_real_july) #TRUE

```

```{r}
#print(ncin_real_tmax)

#LATITUDE SIE ZMNIEJSZA

lat_fore_min <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lat_min")))[2]
lat_fore_max <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lat_max")))[2]
lon_fore_min <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lon_min")))[2]
lon_fore_max <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lon_max")))[2]

lon_fore_start <- which(abs(lon_fore - lon_fore_min) < 0.05)[1]
lon_fore_end <- which(abs(lon_fore - lon_fore_max) < 0.05)[2]
lat_fore_end <- which(abs(lat_fore - lat_fore_min) < 0.05)[2]
lat_fore_start <- which(abs(lat_fore - lat_fore_max) < 0.05)[1]

lat_fore[lat_fore_end]

lon_fore_count <- lon_fore_end - lon_fore_start + 1
lat_fore_count <- lat_fore_end - lat_fore_start + 1

lon_fore <- lon_fore[lon_fore_start:lon_fore_end]
lat_fore <- lat_fore[lat_fore_start:lat_fore_end]


tmax_fore_array <- ncvar_get(ncin_fore_tmax,"air_temperature",start=c(lon_fore_start,lat_fore_start,1),count=c(lon_fore_count,lat_fore_count,length(time_fore)))

#tmax_fore_july <- tmax_fore_array[,,july]

#summary(tmax_fore_july)

#NA values
which(tmax_fore_array == -99999)
#integer(0) <- no need to modify
```

Kelvin to Celsius convertion
```{r}
tmax_fore_array <- kelvin.to.celsius(tmax_fore_array, round = 2)

summary(tmax_fore_array)
summary(tmax_real_array)
```

Visualization
```{r} 
range(na.omit(as.numeric(tmax_fore_array)))
range(na.omit(as.numeric(lon_fore)))
range(na.omit(as.numeric(lat_fore)))

range_fore_lon <- range(na.omit(as.numeric(lon_fore)))[2] - range(na.omit(as.numeric(lon_fore)))[1]
range_fore_lat <- range(na.omit(as.numeric(lat_fore)))[2] - range(na.omit(as.numeric(lat_fore)))[1]

grid_fore <- expand.grid(lon=lon_fore, lat=lat_fore)
head(grid_fore)
cutpts <- c(15, 19, 21, 23, 25, 27, 29, 31, 33, 36)
cutpts

lon<-lon_fore
lat<-lat_fore
levelplot(tmax_fore_array[,,7] ~ lon * lat, data=grid_fore, at=cutpts, cuts=11, pretty=T, col.regions=(rev(brewer.pal(10,"RdBu"))),aspect=range_fore_lat/range_fore_lon)

```

# MIN TEMPERATURE

Data details
```{r}
print(ncin_fore_tmin)

#air_temperature[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

#Checking units
ncatt_get(ncin_fore_tmin,"longitude","units") #degrees_east
ncatt_get(ncin_fore_tmin,"latitude","units") #degrees_north
ncatt_get(ncin_fore_tmin,"time","units") # "days since 2006-1-15 00:00:00"
ncatt_get(ncin_fore_tmin,"air_temperature","units") #K

ncatt_get(ncin_fore_tmax,"air_temperature","long_name") #Daily Minimum Near-Surface Air Temperature
ncatt_get(ncin_fore_tmin,"air_temperature","_FillValue") #-99999
ncatt_get(ncin_fore_tmin,"air_temperature","missing_value") #-99999

```

Extracting variables
```{r}
lon_fore_tmin <- ncvar_get(ncin_fore_tmin, "longitude")
lat_fore_tmin <- ncvar_get(ncin_fore_tmin, "latitude")
time_fore_tmin <- ncvar_get(ncin_fore_tmin, "time")

#clipping data (time)
time_fore_tmin <- time_fore_tmin[1:192]

#clipping data (area)
lat_fore_min <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lat_min")))[2]
lat_fore_max <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lat_max")))[2]
lon_fore_min <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lon_min")))[2]
lon_fore_max <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lon_max")))[2]

lon_fore_start <- which(abs(lon_fore_tmin - lon_fore_min) < 0.05)[1]
lon_fore_end <- which(abs(lon_fore_tmin - lon_fore_max) < 0.05)[2]
lat_fore_end <- which(abs(lat_fore_tmin - lat_fore_min) < 0.05)[2]
lat_fore_start <- which(abs(lat_fore_tmin - lat_fore_max) < 0.05)[1]

lat_fore_tmin[lat_fore_end]

lon_fore_count <- lon_fore_end - lon_fore_start + 1
lat_fore_count <- lat_fore_end - lat_fore_start + 1

lon_fore_tmin <- lon_fore_tmin[lon_fore_start:lon_fore_end]
lat_fore_tmin <- lat_fore_tmin[lat_fore_start:lat_fore_end]

#checking if new variables are necessary
length(lon_fore == lon_fore_tmin) == length(lon_fore_tmin) #TRUE
length(lat_fore == lat_fore_tmin) == length(lat_fore_tmin) #TRUE
length(time_fore == time_fore_tmin) == length(time_fore_tmin) #TRUE
#deleting unnecessary variables
rm(lon_fore_tmin, lat_fore_tmin, time_fore_tmin)


tmin_fore_array <- ncvar_get(ncin_fore_tmin,"air_temperature",start=c(lon_fore_start,lat_fore_start,1),count=c(lon_fore_count,lat_fore_count,length(time_fore)))

#tmin_fore_july <- tmin_fore_array[,,july]
#summary(tmin_fore_july)

#NA values
which(tmin_fore_array == -99999)
#integer(0) <- no need to modify
```


Kelvin to Celsius convertion
```{r}
tmin_fore_array <- kelvin.to.celsius(tmin_fore_array, round = 2)

summary(tmin_fore_array)
summary(tmin_real_array)
```


Visualization
```{r} 
range(na.omit(as.numeric(tmin_fore_array)))
range(na.omit(as.numeric(lon_fore)))
range(na.omit(as.numeric(lat_fore)))


cutpts <- c(4, 6, 8, 10, 12, 14, 16, 18, 20, 22)
cutpts

lon<-lon_fore
lat<-lat_fore
levelplot(tmin_fore_array[,,7] ~ lon * lat, data=grid_fore, at=cutpts, cuts=11, pretty=T, col.regions=(rev(brewer.pal(10,"RdBu"))),aspect=range_fore_lat/range_fore_lon)

```


# PRECIPITATION


Data details
```{r}
print(ncin_fore_ppt)

#precipitation_flux[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

#Checking units
ncatt_get(ncin_fore_ppt,"longitude","units") #degrees_east
ncatt_get(ncin_fore_ppt,"latitude","units") #degrees_north
ncatt_get(ncin_fore_ppt,"time","units") # "days since 2006-01-01 00:00:00"
ncatt_get(ncin_fore_ppt,"precipitation_flux","units") #kg m-2 s-1

ncatt_get(ncin_fore_ppt,"precipitation_flux","long_name") #Precipitation
ncatt_get(ncin_fore_ppt,"precipitation_flux","_FillValue") #-99999
ncatt_get(ncin_fore_ppt,"precipitation_flux","missing_value") #-99999


```
