---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/AGH/INZ/data")
```

#function for running external .Rmd files
```{r}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext=".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output=tmp_file)
  source(file = tmp_file, ...)
}
```

```{r}
source_rmd("../codes/forecast.Rmd")
```

LOADING SAVED WORKSPACE
```{r}
#load("D:/AGH/INZ/codes/manual_save.RData")
#source_rmd("D:/AGH/INZ/codes/shapefiles.Rmd")
#Sys.setlocale("LC_TIME", "C")
```

########################################################################################################

#FITTING GRIDS

COMPARING RESAMPLE CONFIGURATIONS
```{r}

lon_real_raster = raster(ncol=length(lon_real), nrow=length(lat_real))
lat_real_raster = raster(ncol=length(lon_real), nrow=length(lat_real))
lon_fore_raster = raster(ncol=length(lon_fore), nrow=length(lat_fore))
lat_fore_raster = raster(ncol=length(lon_fore), nrow=length(lat_fore))

values(lon_real_raster) = as.vector(grid_real_spei6[,1])
values(lat_real_raster) = as.vector(grid_real_spei6[,2])
values(lon_fore_raster) = as.vector(grid_fore_spei6[,1])
values(lat_fore_raster) = as.vector(grid_fore_spei6[,2])

par(mfrow=c(2,2))
plot(lon_real_raster)                         
plot(lat_real_raster)
plot(lon_fore_raster)
plot(lat_fore_raster)

#######################################################################
#forecast to real (5->4)
lon_fore_bilinear <- resample(lon_fore_raster, lon_real_raster, method='bilinear')
lat_fore_bilinear <- resample(lat_fore_raster, lat_real_raster, method='bilinear')

ncol(lon_real_raster)

mse1=0
for (i in 1:dim(lon_fore_bilinear)[1])
{
  for (j in 1:dim(lon_fore_bilinear)[2])
  {
    print(i)
    mse1 = mse1 + sqrt((lon_fore_bilinear[i,j] - lon_real_raster[i,j])^2 + (lat_fore_bilinear[i,j] - lat_real_raster[i,j])^2)
  }
}
mse1 = mse1 / (dim(lon_fore_bilinear)[1] * dim(lon_fore_bilinear)[2]) 
mse1 #0.01299204


#######################################################################
#real to forecast (4->5)
lon_real_bilinear <- resample(lon_real_raster, lon_fore_raster, method='bilinear')
lat_real_bilinear <- resample(lat_real_raster, lat_fore_raster, method='bilinear')

ncol(lon_real_bilinear)

mse2=0
for (i in 1:dim(lon_real_bilinear)[1])
{
  for (j in 1:dim(lon_real_bilinear)[2])
  {
    print(i)
    mse2 = mse2 + sqrt((lon_real_bilinear[i,j] - lon_fore_raster[i,j])^2 + (lat_real_bilinear[i,j] - lat_fore_raster[i,j])^2)
  }
}
mse2 = mse2 / (dim(lon_real_bilinear)[1] * dim(lon_real_bilinear)[2]) 
mse2 #0.01299202

mse1
mse2
#smaller MSE for second option (real to forecast)
```

PREPARING RASTERSTACKS

SPEI1
```{r}
spei1_real_list = list()
spei1_fore_list = list()

grid_fore_spei1 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei1 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #forecast
  grid_fore_spei1$z<-as.numeric(SPEI1_fore[,,i])
  spei1_fore_raster<-rasterFromXYZ(grid_fore_spei1,crs=4321)
  spei1_fore_list <- append(spei1_fore_list, spei1_fore_raster)
  
  #real
  grid_real_spei1$z<-as.numeric(SPEI1_real[,,i])
  spei1_real_raster<-rasterFromXYZ(grid_real_spei1,crs=4321)
  spei1_real_r2 <- resample(spei1_real_raster, spei1_fore_raster, method='bilinear')
  spei1_real_list <- append(spei1_real_list, spei1_real_r2)
}

real_spei1_stack <- stack(spei1_real_list)
fore_spei1_stack <- stack(spei1_fore_list)
plot(real_spei1_stack)
plot(fore_spei1_stack)
```

SPEI3
```{r}
spei3_real_list = list()
spei3_fore_list = list()

grid_fore_spei3 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei3 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #forecast
  grid_fore_spei3$z<-as.numeric(SPEI3_fore[,,i])
  spei3_fore_raster<-rasterFromXYZ(grid_fore_spei3,crs=4323)
  spei3_fore_list <- append(spei3_fore_list, spei3_fore_raster)
  
  #real
  grid_real_spei3$z<-as.numeric(SPEI3_real[,,i])
  spei3_real_raster<-rasterFromXYZ(grid_real_spei3,crs=4323)
  spei3_real_r2 <- resample(spei3_real_raster, spei3_fore_raster, method='bilinear')
  spei3_real_list <- append(spei3_real_list, spei3_real_r2)
}

real_spei3_stack <- stack(spei3_real_list)
fore_spei3_stack <- stack(spei3_fore_list)
plot(real_spei3_stack)
plot(fore_spei3_stack)
```

SPEI6
```{r}
spei6_real_list = list()
spei6_fore_list = list()

grid_fore_spei6 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei6 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #forecast
  grid_fore_spei6$z<-as.numeric(SPEI6_fore[,,i])
  spei6_fore_raster<-rasterFromXYZ(grid_fore_spei6,crs=4326)
  spei6_fore_list <- append(spei6_fore_list, spei6_fore_raster)
  
  #real
  grid_real_spei6$z<-as.numeric(SPEI6_real[,,i])
  spei6_real_raster<-rasterFromXYZ(grid_real_spei6,crs=4326)
  spei6_real_r2 <- resample(spei6_real_raster, spei6_fore_raster, method='bilinear')
  spei6_real_list <- append(spei6_real_list, spei6_real_r2)
}

real_spei6_stack <- stack(spei6_real_list)
fore_spei6_stack <- stack(spei6_fore_list)
plot(real_spei6_stack)
plot(fore_spei6_stack)

writeRaster(real_spei3_stack, "rasters/stacks/real_spei3_stack.tif", format="GTiff")
writeRaster(fore_spei3_stack, "rasters/stacks/forecast_spei3_stack.tif", format="GTiff")

```

SPEI12
```{r}
spei12_real_list = list()
spei12_fore_list = list()

grid_fore_spei12 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei12 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #forecast
  grid_fore_spei12$z<-as.numeric(SPEI12_fore[,,7])
  spei12_fore_raster<-rasterFromXYZ(grid_fore_spei12,crs=43212)
  spei12_fore_list <- append(spei12_fore_list, spei12_fore_raster)
  
  #real
  grid_real_spei12$z<-as.numeric(SPEI12_real[,,i])
  spei12_real_raster<-rasterFromXYZ(grid_real_spei12,crs=43212)
  spei12_real_r2 <- resample(spei12_real_raster, spei12_fore_raster, method='bilinear')
  spei12_real_list <- append(spei12_real_list, spei12_real_r2)
}

real_spei12_stack <- stack(spei12_real_list)
fore_spei12_stack <- stack(spei12_fore_list)
plot(real_spei12_stack)
plot(fore_spei12_stack)
```

CORRELATION

SPEI1
```{r}
#pearson
corr_p1 <- corLocal(real_spei1_stack, fore_spei1_stack, method="pearson", test=TRUE)
plot(corr_p1)
corr_p1_mask <- mask(corr_p1[[1]], corr_p1[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p1_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s1 <- corLocal(real_spei1_stack, fore_spei1_stack, method="spearman", test=TRUE)
plot(corr_s1)
corr_s1_mask <- mask(corr_s1[[1]], corr_s1[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s1_mask)
plot(both, add=TRUE, lwd=2)
```

SPEI3
```{r}
#pearson
corr_p3 <- corLocal(real_spei3_stack, fore_spei3_stack, method="pearson", test=TRUE)
plot(corr_p3)
corr_p3_mask <- mask(corr_p3[[1]], corr_p3[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p3_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s3 <- corLocal(real_spei3_stack, fore_spei3_stack, method="spearman", test=TRUE)
plot(corr_s3)
corr_s3_mask <- mask(corr_s3[[1]], corr_s3[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s3_mask)
plot(both, add=TRUE, lwd=2)

```

SPEI6
```{r}
#pearson
corr_p6 <- corLocal(real_spei6_stack, fore_spei6_stack, method="pearson", test=TRUE)
plot(corr_p6)
corr_p6_mask <- mask(corr_p6[[1]], corr_p6[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p6_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s6 <- corLocal(real_spei6_stack, fore_spei6_stack, method="spearman", test=TRUE)
plot(corr_s6)
corr_s6_mask <- mask(corr_s6[[1]], corr_s6[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s6_mask)
plot(both, add=TRUE, lwd=2)

```

SPEI12
```{r}
#pearson
corr_p12 <- corLocal(real_spei12_stack, fore_spei12_stack, method="pearson", test=TRUE)
plot(corr_p12)
corr_p12_mask <- mask(corr_p12[[1]], corr_p12[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p12_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s12 <- corLocal(real_spei12_stack, fore_spei12_stack, method="spearman", test=TRUE)
plot(corr_s12)
corr_s12_mask <- mask(corr_s12[[1]], corr_s12[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s12_mask)
plot(both, add=TRUE, lwd=2)

```




Visualization

```{r}
#PEARSON
plot(corr_p1, main="SPEI1")
plot(corr_p3, main="SPEI3")
plot(corr_p6, main="SPEI6")
plot(corr_p12, main="SPEI12")

#masked
par(mfrow=c(2,2))
plot(corr_p1_mask, main = "SPEI1 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p3_mask, main = "SPEI3 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p6_mask, main = "SPEI6 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p12_mask, main = "SPEI12 masked")
plot(both, add=TRUE, lwd=2)


#SPEARMAN
plot(corr_s1, main="SPEI1")
plot(corr_s3, main="SPEI3")
plot(corr_s6, main="SPEI6")
plot(corr_s12, main="SPEI12")

#masked
par(mfrow=c(2,2))
plot(corr_s1_mask, main = "SPEI1 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s3_mask, main = "SPEI3 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s6_mask, main = "SPEI6 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s12_mask, main = "SPEI12 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s12_mask, main = "SPEI12 masked")
plot(both, add=TRUE, lwd=2)
```


##############################################################################################
##############################################################################################


PREPARING RASTERSTACKS - option 2

SPEI1
```{r}
spei1_real_list = list()
spei1_fore_list = list()

grid_fore_spei1 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei1 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{

  #real
  grid_real_spei1$z<-as.numeric(SPEI1_real[,,i])
  spei1_real_raster<-rasterFromXYZ(grid_real_spei1,crs=4321)
  spei1_real_list <- append(spei1_real_list, spei1_real_raster)

  #forecast
  grid_fore_spei1$z<-as.numeric(SPEI1_fore[,,i])
  spei1_fore_raster<-rasterFromXYZ(grid_fore_spei1,crs=4321)
  spei1_fore_r2 <- resample(spei1_fore_raster, spei1_real_raster, method='bilinear')
  spei1_fore_list <- append(spei1_fore_list, spei1_fore_r2)
  
}

real_spei1_stack_2 <- stack(spei1_real_list)
fore_spei1_stack_2 <- stack(spei1_fore_list)
plot(real_spei1_stack_2)
plot(fore_spei1_stack_2)
```

SPEI3
```{r}
spei3_real_list = list()
spei3_fore_list = list()

grid_fore_spei3 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei3 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #real
  grid_real_spei3$z<-as.numeric(SPEI3_real[,,i])
  spei3_real_raster<-rasterFromXYZ(grid_real_spei3,crs=4323)
  spei3_real_list <- append(spei3_real_list, spei3_real_raster)

  #forecast
  grid_fore_spei3$z<-as.numeric(SPEI3_fore[,,i])
  spei3_fore_raster<-rasterFromXYZ(grid_fore_spei3,crs=4323)
  spei3_fore_r2 <- resample(spei3_fore_raster, spei3_real_raster, method='bilinear')
  spei3_fore_list <- append(spei3_fore_list, spei3_fore_r2)
  
  
}

real_spei3_stack_2 <- stack(spei3_real_list)
fore_spei3_stack_2 <- stack(spei3_fore_list)
plot(real_spei3_stack_2)
plot(fore_spei3_stack_2)
```

SPEI6
```{r}
spei6_real_list = list()
spei6_fore_list = list()

grid_fore_spei6 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei6 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #real
  grid_real_spei6$z<-as.numeric(SPEI6_real[,,i])
  spei6_real_raster<-rasterFromXYZ(grid_real_spei6,crs=4326)
  spei6_real_list <- append(spei6_real_list, spei6_real_raster)

  #forecast
  grid_fore_spei6$z<-as.numeric(SPEI6_fore[,,i])
  spei6_fore_raster<-rasterFromXYZ(grid_fore_spei6,crs=4326)
  spei6_fore_r2 <- resample(spei6_fore_raster, spei6_real_raster, method='bilinear')
  spei6_fore_list <- append(spei6_fore_list, spei6_fore_r2)
  
  
}

real_spei6_stack_2 <- stack(spei6_real_list)
fore_spei6_stack_2 <- stack(spei6_fore_list)
plot(real_spei6_stack_2)
plot(fore_spei6_stack_2)
```

SPEI12
```{r}
spei12_real_list = list()
spei12_fore_list = list()

grid_fore_spei12 <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_spei12 <- expand.grid(lon=lon_real, lat=lat_real)

for (i in july_real) #same as july_fore
{
  #real
  grid_real_spei12$z<-as.numeric(SPEI12_real[,,i])
  spei12_real_raster<-rasterFromXYZ(grid_real_spei12,crs=43212)
  spei12_real_list <- append(spei12_real_list, spei12_real_raster)

  #forecast
  grid_fore_spei12$z<-as.numeric(SPEI12_fore[,,i])
  spei12_fore_raster<-rasterFromXYZ(grid_fore_spei12,crs=43212)
  spei12_fore_r2 <- resample(spei12_fore_raster, spei12_real_raster, method='bilinear')
  spei12_fore_list <- append(spei12_fore_list, spei12_fore_r2)
  
  
}

real_spei12_stack_2 <- stack(spei12_real_list)
fore_spei12_stack_2 <- stack(spei12_fore_list)
plot(real_spei12_stack_2)
plot(fore_spei12_stack_2)
```



CORRELATION - option 2

SPEI1
```{r}
#pearson
corr_p1 <- corLocal(real_spei1_stack_2, fore_spei1_stack_2, method="pearson", test=TRUE)
plot(corr_p1)
corr_p1_mask <- mask(corr_p1[[1]], corr_p1[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p1_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s1 <- corLocal(real_spei1_stack_2, fore_spei1_stack_2, method="spearman", test=TRUE)
plot(corr_s1)
corr_s1_mask <- mask(corr_s1[[1]], corr_s1[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s1_mask)
plot(both, add=TRUE, lwd=2)
```

SPEI3
```{r}
#pearson
corr_p3 <- corLocal(real_spei3_stack_2, fore_spei3_stack_2, method="pearson", test=TRUE)
plot(corr_p3)
corr_p3_mask <- mask(corr_p3[[1]], corr_p3[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p3_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s3 <- corLocal(real_spei3_stack_2, fore_spei3_stack_2, method="spearman", test=TRUE)
plot(corr_s3)
corr_s3_mask <- mask(corr_s3[[1]], corr_s3[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s3_mask)
plot(both, add=TRUE, lwd=2)

```

SPEI6
```{r}
#pearson
corr_p6 <- corLocal(real_spei6_stack_2, fore_spei6_stack_2, method="pearson", test=TRUE)
plot(corr_p6)
corr_p6_mask <- mask(corr_p6[[1]], corr_p6[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p6_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s6 <- corLocal(real_spei6_stack_2, fore_spei6_stack_2, method="spearman", test=TRUE)
plot(corr_s6)
corr_s6_mask <- mask(corr_s6[[1]], corr_s6[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s6_mask)
plot(both, add=TRUE, lwd=2)

```

SPEI12
```{r}
#pearson
corr_p12 <- corLocal(real_spei12_stack_2, fore_spei12_stack_2, method="pearson", test=TRUE)
plot(corr_p12)
corr_p12_mask <- mask(corr_p12[[1]], corr_p12[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p12_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s12 <- corLocal(real_spei12_stack_2, fore_spei12_stack_2, method="spearman", test=TRUE)
plot(corr_s12)
corr_s12_mask <- mask(corr_s12[[1]], corr_s12[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s12_mask)
plot(both, add=TRUE, lwd=2)

```




Visualization

```{r}
#PEARSON
plot(corr_p1, main="SPEI1")
plot(corr_p3, main="SPEI3")
plot(corr_p6, main="SPEI6")
plot(corr_p12, main="SPEI12")

#masked
par(mfrow=c(2,2))
plot(corr_p1_mask, main = "SPEI1 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p3_mask, main = "SPEI3 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p6_mask, main = "SPEI6 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_p12_mask, main = "SPEI12 masked")
plot(both, add=TRUE, lwd=2)


#SPEARMAN
plot(corr_s1, main="SPEI1")
plot(corr_s3, main="SPEI3")
plot(corr_s6, main="SPEI6")
plot(corr_s12, main="SPEI12")

#masked
par(mfrow=c(2,2))
plot(corr_s1_mask, main = "SPEI1 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s3_mask, main = "SPEI3 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s6_mask, main = "SPEI6 masked")
plot(both, add=TRUE, lwd=2)
plot(corr_s12_mask, main = "SPEI12 masked")
plot(both, add=TRUE, lwd=2)
```


CORRELATION OF OTHER VARIABLES (tmax, tmin, ppt)


PREPARING RASTERSTACKS

TMAX 
```{r}
tmax_real_list = list()
tmax_fore_list = list()

grid_fore_tmax <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_tmax <- expand.grid(lon=lon_real, lat=lat_real)

for (i in seq(1, length(time_fore))) #same as july_fore
{
  #forecast
  grid_fore_tmax$z<-as.numeric(tmax_fore_array[,,i])
  tmax_fore_raster<-rasterFromXYZ(grid_fore_tmax,crs=4321)
  tmax_fore_list <- append(tmax_fore_list, tmax_fore_raster)
  
  #real
  grid_real_tmax$z<-as.numeric(tmax_real_array[,,i])
  tmax_real_raster<-rasterFromXYZ(grid_real_tmax,crs=4321)
  tmax_real_r2 <- resample(tmax_real_raster, tmax_fore_raster, method='bilinear')
  tmax_real_list <- append(tmax_real_list, tmax_real_r2)
}

real_tmax_stack <- stack(tmax_real_list)
fore_tmax_stack <- stack(tmax_fore_list)
plot(real_tmax_stack)
plot(fore_tmax_stack)
```

TMIN
```{r}
tmin_real_list = list()
tmin_fore_list = list()

grid_fore_tmin <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_tmin <- expand.grid(lon=lon_real, lat=lat_real)

for (i in seq(1, length(time_fore))) #same as july_fore
{
  #forecast
  grid_fore_tmin$z<-as.numeric(tmin_fore_array[,,i])
  tmin_fore_raster<-rasterFromXYZ(grid_fore_tmin,crs=4321)
  tmin_fore_list <- append(tmin_fore_list, tmin_fore_raster)
  
  #real
  grid_real_tmin$z<-as.numeric(tmin_real_array[,,i])
  tmin_real_raster<-rasterFromXYZ(grid_real_tmin,crs=4321)
  tmin_real_r2 <- resample(tmin_real_raster, tmin_fore_raster, method='bilinear')
  tmin_real_list <- append(tmin_real_list, tmin_real_r2)
}

real_tmin_stack <- stack(tmin_real_list)
fore_tmin_stack <- stack(tmin_fore_list)
plot(real_tmin_stack)
plot(fore_tmin_stack)
```

PPT
```{r}
ppt_real_list = list()
ppt_fore_list = list()

grid_fore_ppt <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_real_ppt <- expand.grid(lon=lon_real, lat=lat_real)

for (i in seq(1, length(time_fore))) #same as july_fore
{
  #forecast
  grid_fore_ppt$z<-as.numeric(ppt_fore_array[,,i])
  ppt_fore_raster<-rasterFromXYZ(grid_fore_ppt,crs=4321)
  ppt_fore_list <- append(ppt_fore_list, ppt_fore_raster)
  
  #real
  grid_real_ppt$z<-as.numeric(ppt_real_array[,,i])
  ppt_real_raster<-rasterFromXYZ(grid_real_ppt,crs=4321)
  ppt_real_r2 <- resample(ppt_real_raster, ppt_fore_raster, method='bilinear')
  ppt_real_list <- append(ppt_real_list, ppt_real_r2)
}

real_ppt_stack <- stack(ppt_real_list)
fore_ppt_stack <- stack(ppt_fore_list)
plot(real_ppt_stack)
plot(fore_ppt_stack)
```

CORRELATION

TMAX
```{r}
#pearson
corr_p_tmax <- corLocal(real_tmax_stack, fore_tmax_stack, method="pearson", test=TRUE)
plot(corr_p_tmax)
corr_p_tmax_mask <- mask(corr_p_tmax[[1]], corr_p_tmax[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p_tmax_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s_tmax <- corLocal(real_tmax_stack, fore_tmax_stack, method="spearman", test=TRUE)
plot(corr_s_tmax)
corr_s_tmax_mask <- mask(corr_s_tmax[[1]], corr_s_tmax[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s_tmax_mask)
plot(both, add=TRUE, lwd=2)
```

TMIN
```{r}
#pearson
corr_p_tmin <- corLocal(real_tmin_stack, fore_tmin_stack, method="pearson", test=TRUE)
plot(corr_p_tmin)
corr_p_tmin_mask <- mask(corr_p_tmin[[1]], corr_p_tmin[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p_tmin_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s_tmin <- corLocal(real_tmin_stack, fore_tmin_stack, method="spearman", test=TRUE)
plot(corr_s_tmin)
corr_s_tmin_mask <- mask(corr_s_tmin[[1]], corr_s_tmin[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s_tmin_mask)
plot(both, add=TRUE, lwd=2)
```

PPT
```{r}
#pearson
corr_p_ppt <- corLocal(real_ppt_stack, fore_ppt_stack, method="pearson", test=TRUE)
plot(corr_p_ppt)
corr_p_ppt_mask <- mask(corr_p_ppt[[1]], corr_p_ppt[[2]] < 0.05, maskvalue=FALSE)
plot(corr_p_ppt_mask)
plot(both, add=TRUE, lwd=2)

#spearman
corr_s_ppt <- corLocal(real_ppt_stack, fore_ppt_stack, method="spearman", test=TRUE)
plot(corr_s_ppt)
corr_s_ppt_mask <- mask(corr_s_ppt[[1]], corr_s_ppt[[2]] < 0.05, maskvalue=FALSE)
plot(corr_s_ppt_mask)
plot(both, add=TRUE, lwd=2)
```