---
title: "R Notebook"
output: html_notebook
---

USAGE: variables generated with "kod_real.Rmd" script are needed to run this file properly

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/AGH/INZ/data")
```

#function for running external .Rmd files
```{r}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext=".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output=tmp_file)
  source(file = tmp_file, ...)
}
```


RUNNING NECESSARY FILES (kod_real.Rmd, kod_shapefiles.Rmd, setup.R)
```{r}
source_rmd("../codes/real.Rmd")

```
LOADING SAVED WORKSPACE
```{r}
#load("D:/AGH/INZ/codes/manual_save.RData")
#source_rmd("D:/AGH/INZ/codes/shapefiles.Rmd")
#Sys.setlocale("LC_TIME", "C")
```


##################################################################################################################################################

# FORECAST

Loading data
```{r}
#setwd("D:/AGH/INZ/data")

ncin_fore_ppt<-nc_open("forecast/CHELSAcmip5ts_pr_ACCESS1-3_rcp45_2006-2029_V1.1.nc")
ncin_fore_tmax<-nc_open("forecast/CHELSAcmip5ts_tasmax_ACCESS1-3_rcp45_2006-2029_V1.1.nc")
ncin_fore_tmin<-nc_open("forecast/CHELSAcmip5ts_tasmin_ACCESS1-3_rcp45_2006-2029_V1.1.nc")

```

# MAX TEMPERATURE

Data details
```{r}
#print(ncin_fore_tmax)

#air_temperature[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

ncatt_get(ncin_fore_tmax, 0, "Title")
ncatt_get(ncin_fore_tmax, 0, "References")
ncatt_get(ncin_fore_tmax, 0, "history")
ncatt_get(ncin_fore_tmax, 0, "Conventions")

#Checking units
ncatt_get(ncin_fore_tmax,"longitude","units") #degrees_east
ncatt_get(ncin_fore_tmax,"latitude","units") #degrees_north
ncatt_get(ncin_fore_tmax,"time","units") # "days since 2006-1-15 00:00:00"
ncatt_get(ncin_fore_tmax,"air_temperature","units") #K

ncatt_get(ncin_fore_tmax,"air_temperature","long_name") #Daily Maximum Near-Surface Air Temperature
ncatt_get(ncin_fore_tmax,"air_temperature","_FillValue") #-99999
ncatt_get(ncin_fore_tmax,"air_temperature","missing_value") #-99999


```



Extracting variables
```{r}
lon_fore <- ncvar_get(ncin_fore_tmax, "longitude")
lat_fore <- ncvar_get(ncin_fore_tmax, "latitude")
time_fore <- ncvar_get(ncin_fore_tmax, "time")

dim(lon_fore) #8017
dim(lat_fore) #3875
dim(time_fore) # 288 -> 24 years, 12 months each

head(lon_fore) # - => +
head(lat_fore) # + => -
head(time_fore)

```


Time units correction
```{r}
time_units <- ncatt_get(ncin_fore_tmax,"time","units")
time_units <- strsplit(time_units$value, " ")
time_units <- strsplit(unlist(time_units)[3], "-")

time_year <- as.integer(unlist(time_units)[1])
time_month <- as.integer(unlist(time_units)[2])
time_day <- as.integer(unlist(time_units)[3])

time_fore <- chron(time_fore, origin = c(time_month, time_day, time_year))
Sys.setlocale("LC_TIME", "C")
time_fore <- as.yearmon(time_fore)

head(time_fore)
tail(time_fore)
#time range: Jan 2006 - Dec 2029

#limit to range Jan 2006 - Dec 2021
time_fore <- time_fore[1:192]
#which(time_fore == tail(time_fore, n=1))

```

#LIMITING FORECAST DATA TO THE RANGE OF REAL DATA
```{r}
#LATITUDE IS DECREASING

lat_fore_min <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lat_min")))[2]
lat_fore_max <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lat_max")))[2]
lon_fore_min <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lon_min")))[2]
lon_fore_max <- (as.numeric(ncatt_get(ncin_real_tmax, 0,"geospatial_lon_max")))[2]

#finding nearest gridpoint
lon_fore_start <- which(abs(lon_fore - lon_fore_min) < 0.02)[1]
lon_fore_end <- which(abs(lon_fore - lon_fore_max) < 0.02)[1]
lat_fore_end <- which(abs(lat_fore - lat_fore_min) < 0.02)[1]
lat_fore_start <- which(abs(lat_fore - lat_fore_max) < 0.023)[1]

lat_fore[lat_fore_end]

#gripoints count
lon_fore_count <- lon_fore_end - lon_fore_start + 1
lat_fore_count <- lat_fore_end - lat_fore_start + 1

lon_fore <- lon_fore[lon_fore_start:lon_fore_end]
lat_fore <- lat_fore[lat_fore_start:lat_fore_end]


tmax_fore_array <- ncvar_get(ncin_fore_tmax,"air_temperature",start=c(lon_fore_start,lat_fore_start,1),count=c(lon_fore_count,lat_fore_count,length(time_fore)))

#NA values
which(tmax_fore_array == -99999)
#integer(0) <- no need to modify
```

Kelvin to Celsius convertion
```{r}
tmax_fore_array <- kelvin.to.celsius(tmax_fore_array, round = 2)

summary(tmax_fore_array)
summary(tmax_real_array)
```

Visualization
```{r} 
range(na.omit(as.numeric(tmax_fore_array[,,7])))
range(na.omit(as.numeric(lon_fore)))
range(na.omit(as.numeric(lat_fore)))

range_fore_lon <- range(na.omit(as.numeric(lon_fore)))[2] - range(na.omit(as.numeric(lon_fore)))[1]
range_fore_lat <- range(na.omit(as.numeric(lat_fore)))[2] - range(na.omit(as.numeric(lat_fore)))[1]

grid_fore <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_fore$z<-as.numeric(tmax_fore_array[,,7])

#cutpts <- c(15, 19, 21, 23, 25, 27, 29, 31, 33, 36)
cuts<-seq(14,36,2)

levelplot(z ~ lon * lat, data=grid_fore, at=cuts, pretty=T, col.regions=(rev(brewer.pal(11,"RdBu"))),aspect=range_fore_lat/range_fore_lon, main = "Forecasted maximum temperature - July 2006")

```

# MIN TEMPERATURE

Data details
```{r}
print(ncin_fore_tmin)

#air_temperature[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

#Checking units
ncatt_get(ncin_fore_tmin,"longitude","units") #degrees_east
ncatt_get(ncin_fore_tmin,"latitude","units") #degrees_north
ncatt_get(ncin_fore_tmin,"time","units") # "days since 2006-1-15 00:00:00"
ncatt_get(ncin_fore_tmin,"air_temperature","units") #K

ncatt_get(ncin_fore_tmax,"air_temperature","long_name") #Daily Minimum Near-Surface Air Temperature
ncatt_get(ncin_fore_tmin,"air_temperature","_FillValue") #-99999
ncatt_get(ncin_fore_tmin,"air_temperature","missing_value") #-99999

```

Extracting variables
```{r}
lon_fore_tmin <- ncvar_get(ncin_fore_tmin, "longitude")
lat_fore_tmin <- ncvar_get(ncin_fore_tmin, "latitude")
time_fore_tmin <- ncvar_get(ncin_fore_tmin, "time")

#clipping data (time)
time_fore_tmin <- time_fore_tmin[1:192]
head(time_fore_tmin)

time_units <- ncatt_get(ncin_fore_tmin,"time","units")
time_units <- strsplit(time_units$value, " ")
time_units <- strsplit(unlist(time_units)[3], "-")

time_year <- as.integer(unlist(time_units)[1])
time_month <- as.integer(unlist(time_units)[2])
time_day <- as.integer(unlist(time_units)[3])

time_fore_tmin <- chron(time_fore_tmin, origin = c(time_month, time_day, time_year))
Sys.setlocale("LC_TIME", "C")
time_fore_tmin <- as.yearmon(time_fore_tmin)

head(time_fore_tmin)
tail(time_fore_tmin)
#time range: Jan 2006 - Dec 2029

#limit to range Jan 2006 - Dec 2021
time_fore_tmin <- time_fore_tmin[1:192]

#clipping data (area)
lat_fore_min <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lat_min")))[2]
lat_fore_max <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lat_max")))[2]
lon_fore_min <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lon_min")))[2]
lon_fore_max <- (as.numeric(ncatt_get(ncin_real_tmin, 0,"geospatial_lon_max")))[2]

lon_fore_start <- which(abs(lon_fore_tmin - lon_fore_min) < 0.02)[1]
lon_fore_end <- which(abs(lon_fore_tmin - lon_fore_max) < 0.02)[1]
lat_fore_end <- which(abs(lat_fore_tmin - lat_fore_min) < 0.02)[1]
lat_fore_start <- which(abs(lat_fore_tmin - lat_fore_max) < 0.023)[1]

lat_fore_tmin[lat_fore_end]

lon_fore_count <- lon_fore_end - lon_fore_start + 1
lat_fore_count <- lat_fore_end - lat_fore_start + 1

lon_fore_tmin <- lon_fore_tmin[lon_fore_start:lon_fore_end]
lat_fore_tmin <- lat_fore_tmin[lat_fore_start:lat_fore_end]

#checking if new variables are necessary
which((lon_fore == lon_fore_tmin) == FALSE) #0
which((lat_fore == lat_fore_tmin) == FALSE) #0
which((time_fore == time_fore_tmin) == FALSE) #0
#deleting unnecessary variables
rm(lon_fore_tmin, lat_fore_tmin, time_fore_tmin)


tmin_fore_array <- ncvar_get(ncin_fore_tmin,"air_temperature",start=c(lon_fore_start,lat_fore_start,1),count=c(lon_fore_count,lat_fore_count,length(time_fore)))

#NA values
which(tmin_fore_array == -99999)
#integer(0) <- no need to modify
```


Kelvin to Celsius convertion
```{r}
tmin_fore_array <- kelvin.to.celsius(tmin_fore_array, round = 2)

summary(tmin_fore_array)
summary(tmin_real_array)
```


Visualization
```{r} 
range(na.omit(as.numeric(tmin_fore_array[,,7])))
range(na.omit(as.numeric(lon_fore)))
range(na.omit(as.numeric(lat_fore)))

grid_fore <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_fore$z<-as.numeric(tmin_fore_array[,,7])

cuts<-seq(6,22,2)

levelplot(z ~ lon * lat, data=grid_fore, at=cuts, pretty=T, col.regions=(rev(brewer.pal(11,"RdBu"))),aspect=range_fore_lat/range_fore_lon, main = "Forecasted minimum temperature - July 2006")
```


# PRECIPITATION

Data details
```{r}
print(ncin_fore_ppt)

#precipitation_flux[longitude,latitude,time]
#coordinate_system: WGS84,EPSG:4326

#Checking units
ncatt_get(ncin_fore_ppt,"longitude","units") #degrees_east
ncatt_get(ncin_fore_ppt,"latitude","units") #degrees_north
ncatt_get(ncin_fore_ppt,"time","units") # "days since 2006-01-01 00:00:00"
ncatt_get(ncin_fore_ppt,"precipitation_flux","units") #kg m-2 s-1

ncatt_get(ncin_fore_ppt,"precipitation_flux","long_name") #Precipitation
ncatt_get(ncin_fore_ppt,"precipitation_flux","_FillValue") #-99999
ncatt_get(ncin_fore_ppt,"precipitation_flux","missing_value") #-99999


```

Extracting variables
```{r}
lon_fore_ppt <- ncvar_get(ncin_fore_ppt, "longitude")
lat_fore_ppt <- ncvar_get(ncin_fore_ppt, "latitude")
time_fore_ppt <- ncvar_get(ncin_fore_ppt, "time")

```


```{r}
time_units <- ncatt_get(ncin_fore_ppt,"time","units")
time_units <- strsplit(time_units$value, " ")
time_units <- strsplit(unlist(time_units)[3], "-")

time_year <- as.integer(unlist(time_units)[1])
time_month <- as.integer(unlist(time_units)[2])
time_day <- as.integer(unlist(time_units)[3])

time_fore_ppt <- chron(time_fore_ppt, origin = c(time_month, time_day, time_year))
Sys.setlocale("LC_TIME", "C")
time_fore_ppt <- as.yearmon(time_fore_ppt)

head(time_fore_ppt)
tail(time_fore_ppt)
#time range: Jan 2006 - Dec 2029

head(time_fore)
tail(time_fore)
#limit to range Jan 2006 - Dec 2021
time_fore_ppt <- time_fore_ppt[1:192]


#clipping data (area)
lat_fore_min <- (as.numeric(ncatt_get(ncin_real_ppt, 0,"geospatial_lat_min")))[2]
lat_fore_max <- (as.numeric(ncatt_get(ncin_real_ppt, 0,"geospatial_lat_max")))[2]
lon_fore_min <- (as.numeric(ncatt_get(ncin_real_ppt, 0,"geospatial_lon_min")))[2]
lon_fore_max <- (as.numeric(ncatt_get(ncin_real_ppt, 0,"geospatial_lon_max")))[2]

lon_fore_start <- which(abs(lon_fore_ppt - lon_fore_min) < 0.05)[1]
lon_fore_end <- which(abs(lon_fore_ppt - lon_fore_max) < 0.05)[2]
lat_fore_end <- which(abs(lat_fore_ppt - lat_fore_min) < 0.05)[2]
lat_fore_start <- which(abs(lat_fore_ppt - lat_fore_max) < 0.05)[1]

lat_fore_ppt[lat_fore_end]

lon_fore_count <- lon_fore_end - lon_fore_start + 1
lat_fore_count <- lat_fore_end - lat_fore_start + 1

lon_fore_ppt <- lon_fore_ppt[lon_fore_start:lon_fore_end]
lat_fore_ppt <- lat_fore_ppt[lat_fore_start:lat_fore_end]

#checking if new variables are necessary
which((lon_fore == lon_fore_ppt) == FALSE) #0
which((lat_fore == lat_fore_ppt) == FALSE) #0
which((time_fore == time_fore_ppt) == FALSE) #0
#deleting unnecessary variables
rm(lon_fore_ppt, lat_fore_ppt, time_fore_ppt)


ppt_fore_array <- ncvar_get(ncin_fore_ppt,"precipitation_flux",start=c(lon_fore_start,lat_fore_start,1),count=c(lon_fore_count,lat_fore_count,length(time_fore)))


#NA values
which(ppt_fore_array == -99999)
#integer(0) <- no need to modify
```


kg m-2 s-1 to mm convertion
```{r}
sec_per_day <- 86400

for (i in 1 : length(time_fore)) {
  ppt_fore_array[,,i] <- ppt_fore_array[,,i] * sec_per_day * days_in_month(time_fore[i])
}

summary(ppt_fore_array)
summary(ppt_real_array)
```


Visualization
```{r} 
range(na.omit(as.numeric(ppt_fore_array[,,7])))
range(na.omit(as.numeric(lon_fore)))
range(na.omit(as.numeric(lat_fore)))


grid_fore <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_fore$z<-as.numeric(ppt_fore_array[,,7])

cuts<-seq(00,100,10)

levelplot(z ~ lon * lat, data=grid_fore, at=cuts, pretty=T, col.regions=(rev(brewer.pal(11,"RdBu"))),aspect=range_fore_lat/range_fore_lon, main = "Forecasted precipitation - July 2006")

```

PET
```{r}
tmax_fore_dim <- dim(tmax_fore_array)
PET_fore <- array(rep(NA, prod(tmax_fore_dim)), dim = tmax_fore_dim)

for (i in 1 : tmax_fore_dim[1])
{
  print(i)
  for (j in 1 : tmax_fore_dim[2])
  {
      PET_fore[i,j,] <- hargreaves(tmin_fore_array[i,j,], tmax_fore_array[i,j,], lat=lat_fore[j], na.rm=TRUE)
  }
}


range(na.omit(as.numeric(PET_fore[,,7])))
median(PET_fore[,,7])

grid_fore_pet <- expand.grid(lon=lon_fore, lat=lat_fore)
grid_fore_pet$z<-as.numeric(PET_fore[,,7])

cuts<-c(90, seq(130,210,20))

levelplot(z ~ lon * lat, data=grid_fore_pet, at=cuts, pretty=T, col.regions=(rev(brewer.pal(6,"RdBu"))),aspect=range_fore_lat/range_fore_lon, main = "PET - July 2006")

```

SPEI
```{r}

SPEI1_fore <- array(rep(NA, prod(tmax_fore_dim)), dim = tmax_fore_dim)

for (i in 1:tmax_fore_dim[1]){
  print(i)
	for (j in 1:tmax_fore_dim[2]){
		a<-as.vector(spei( as.numeric(ppt_fore_array[i,j,]) - as.numeric(PET_fore[i,j,]), 1, na.rm = TRUE)) #1 month
		SPEI1_fore[i,j,]<-as.vector(a$fitted)		
	}
}

SPEI3_fore <- array(rep(NA, prod(tmax_fore_dim)), dim = tmax_fore_dim)

for (i in 1:tmax_fore_dim[1]){
  print(i)
	for (j in 1:tmax_fore_dim[2]){
		a<-as.vector(spei( as.numeric(ppt_fore_array[i,j,]) - as.numeric(PET_fore[i,j,]), 3, na.rm = TRUE)) #3 months
		SPEI3_fore[i,j,]<-as.vector(a$fitted)		
	}
}

SPEI6_fore <- array(rep(NA, prod(tmax_fore_dim)), dim = tmax_fore_dim)

for (i in 1:tmax_fore_dim[1]){
  print(i)
	for (j in 1:tmax_fore_dim[2]){
		a<-as.vector(spei( as.numeric(ppt_fore_array[i,j,]) - as.numeric(PET_fore[i,j,]), 6, na.rm = TRUE)) #6 months
		SPEI6_fore[i,j,]<-as.vector(a$fitted)		
	}
}


SPEI12_fore <- array(rep(NA, prod(tmax_fore_dim)), dim = tmax_fore_dim)

for (i in 1:tmax_fore_dim[1]){
  print(i)
	for (j in 1:tmax_fore_dim[2]){
		a<-as.vector(spei( as.numeric(ppt_fore_array[i,j,]) - as.numeric(PET_fore[i,j,]), 12, na.rm = TRUE)) #12 months
		SPEI12_fore[i,j,]<-as.vector(a$fitted)		
	}
}


```

Visualization
```{r}
range(na.omit(as.numeric(SPEI6_fore[,,7])))
pal <- colorRampPalette(c("red","blue"))
cuts<-seq(-3,3,0.5)

#indexes of every july
july_fore <- seq(7, length(time_fore), by=12)
july <- july_fore
for (i in july_fore)
{
  
  grid_fore_spei1 <- expand.grid(lon=lon_fore, lat=lat_fore)
  grid_fore_spei1$z<-as.numeric(SPEI1_fore[,,19])
  
  levelplot(z ~ lon * lat, data=grid_fore_spei1, main = "Forecasted SPEI1 - July 2007")
  

  spei1_fore_raster<-rasterFromXYZ(grid_fore_spei1,crs=4326)
  
  #writeRaster(spei6_fore_raster, paste("./rasters/forecast_", unlist(strsplit(toString(time_fore[i]), " "))[2], ".tif", sep=''), format="GTiff")
  
  plot(spei12_fore_raster,main=paste("6-month SPEI based on forecasted data - ", time_real[i]),col=pal(12),breaks=cuts)
  plot(both,add=TRUE,lwd=2)
  
  ## crop and mask
  #ex <- extent(as_Spatial(both))
  #spei6_fore_r_cropped <- crop(spei6_fore_raster, ex)
  #spei6_fore_r_masked <- mask(spei6_fore_r_cropped, as_Spatial(both))

  #plot(spei6_fore_r_masked, main=paste("6-month SPEI based on forecasted data - ", time_real[i]), col=pal(12), breaks=cuts)
  #plot(as_Spatial(both), add=TRUE, lwd=2)
}
```


```{r}
save(list=c("both", "north", "south", "pol", "july", "lat_fore", "lon_fore", "lat_real", "lon_real", "PET_fore", "PET_real", "ppt_fore_array", "ppt_real_array", "SPEI1_fore", "SPEI1_real", "SPEI3_fore", "SPEI3_real", "SPEI6_fore", "SPEI6_real", "SPEI12_fore", "SPEI12_real", "time_fore", "time_real", "tmax_fore_array", "tmax_real_array", "tmin_fore_array", "tmin_real_array"), file="manual_real_fore+full.RData")
```